The five pillars of computational reproducibility: Bioinformatics and beyond

### 2023-05-30 manuscript/5pillars.Rmd
adding collabs
link decay addressed #1

Without data sharing, computational research is not reproducible nor auditable.
Lack of data sharing is one of the key reasons why research is irreproducible
[@Ioannidis2009-hv, @Archmiller2020-er ].
Data sharing is also one of the key features of "open science", which is characterised as science
that is collaborative, transparent, accessible and inclusive [@noauthor_2023-uq].
Indeed, inspecting raw data can uncover research integrity problems [@Miyakawa2020-yj],
In addition to facilitating reproduction and auditing, sharing enables reuse in other contexts.
Data reuse increases efficiency, as it prevents redundant research expenditure and facilitates new
research ideas that were previously impossible [@Wilkinson2016-qq].
In terms of research rigour, inspecting raw data can uncover research integrity problems [@Miyakawa2020-yj],
In light of this, a research article without supporting data and code is much like a press release or
advertisement where the claims made cannot be verified as true or not.

@@

2. Avoid commodity cloud storage as these are impermanent.

@@

4. Provide detailed metadata.

5. Preferably archive and share raw data and use existing standards for the discipline.

6. Use file formats that are compatible with many different types of software; e.g.: comma-separated
values (CSV) files are preferred over Excel files.

@@

* Pierre Poulain @pierrepo

>>>>

Without data sharing, computational research is not reproducible nor auditable.
Lack of data sharing is one of the key reasons why research is irreproducible
[@Ioannidis2009-hv, @Archmiller2020-er ].
Data sharing is also one of the key features of "open science", which is characterised as science
that is collaborative, transparent, accessible and inclusive [@noauthor_2023-uq].
In addition to facilitating reproduction and auditing, sharing enables reuse in other contexts.
Data reuse increases efficiency, as it prevents redundant research expenditure and facilitates new
research ideas that were previously impossible [@Wilkinson2016-qq].
In terms of research rigour, inspecting raw data can uncover research integrity problems [@Miyakawa2020-yj],
In light of this, a research article without supporting data and code is much like a press release or
advertisement where the claims made cannot be verified as true or not.

@@

2. Avoid commodity cloud storage as these are impermanent and susceptible to link decay @Hennessey2013-tj.

@@

4. Preferably archive and share raw data and use existing standards for the discipline.

5. Use file formats that are compatible with many different types of software; e.g.: comma-separated
values (CSV) files are preferred over Excel files.

6. Provide detailed metadata; e.g.: sample descriptions that match the manuscript; describe the columns
in tabular data.

@@

* Pierre Poulain @pierrepo (elevate to author)
